Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2018-09-26 Documents/M1/ornlReport/endeve_etal_2017_ORNL_TM_501.tex
ornl report on transport

\documentclass[10pt,preprint]{aastex}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{cite}
\usepackage{times}
\usepackage{url}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{yhmath}
\usepackage{natbib}
\usepackage{../definitions}
\hypersetup{
  bookmarksnumbered = true,
  bookmarksopen=false,
  pdfborder=0 0 0,         % make all links invisible, so the pdf looks good when printed
  pdffitwindow=true,      % window fit to page when opened
  pdfnewwindow=true, % links in new window
  colorlinks=true,           % false: boxed links; true: colored links
  linkcolor=blue,            % color of internal links
  citecolor=magenta,    % color of links to bibliography
  filecolor=magenta,     % color of file links
  urlcolor=cyan              % color of external links
}

\newcommand{\ee}[1]{{\color{red} #1}}
\newcommand{\dx}{\Delta x}
\newcommand{\pbK}{\partial\mathbf{K}}
\newcommand{\UDG}{\bU_{\mbox{\tiny DG}}}
\newcommand{\sumx}{\sum_{i=1}^{d}}
\newcommand{\NES}{\mbox{\tiny NES}}

@@

\title{Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport}
\author{Eirik Endeve\altaffilmark{1}, et al.}
\altaffiltext{1}{Computational and Applied Mathematics Group, Oak Ridge National Laboratory, Oak Ridge, TN 37831-6354, USA; endevee@ornl.gov}

\begin{abstract}
We present a nodal Discontinuous Galerkin (DG) method for solving the non-relativistic two-moment model for neutrino transport.  
The method --- ultimately targeted for simulations of core-collapse supernovae --- includes the use of curvilinear coordinates and non-ideal microphysics (nuclear equation of state and weak interactions) for neutrino-matter interactions.  
\end{abstract}

\tableofcontents

\section{Introduction}

We will use inits in which the speed of light is unity ($c=1$).

\section{Two-Moment Model for Neutrino Transport}
\label{sec:twoMomentModel}

\subsection{Basic Mathematical Model}

For general curvilinear spatial coordinates $\{x^{i}\}$, the two-moment model for radiation transport takes the form \citep[see, e.g.,][for fully general relativistic treatments]{shibata_etal_2011,cardall_etal_2013a}
\begin{align}
  &\f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\cJ\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\cH^{i}\,\Big)
  =\tilde{\chi}\,\big(\cJ_{0}-\cJ\big) \nonumber \\
  &\hspace{12pt}
  +\big(4\pi-\cJ\big)\int_{\bbR^{+}}R^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  -\cJ\int_{\bbR^{+}}R^{\Out}(\epsilonNu,\epsilonNu')\,\big(4\pi-\cJ(\epsilonNu')\big)\,dV_{\epsilonNu'},
  \label{eq:momentEquationEnergy}
\end{align}
\begin{align}
  &\f{1}{\sqrt{\gamma}}\pderiv{}{t}\Big(\,\sqrt{\gamma}\,\cH_{j}\,\Big)
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\cK^{i}_{~j}\,\Big)
  =\f{1}{2}\,\cK^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}-\kappa\,\cH_{j} \nonumber \\
  &\hspace{12pt}
  -\cH_{j}\int_{\bbR^{+}}\big[\,R^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')+R^{\Out}(\epsilonNu,\epsilonNu')\big(4\pi-\cJ(\epsilonNu')\big)\,\big]\,dV_{\epsilonNu'},
  \label{eq:momentEquationMomentum}
\end{align}
where $\cJ$, $\cH^{i}$, and $\cK^{ij}$ are defined as angular moments of a positive neutrino distribution function $f(\omegaNu,\epsilonNu,\bx,t)$
\begin{equation}
  \big\{\,\cJ,\,\cH^{i},\,\cK^{ij}\,\big\}(\epsilonNu,\bx,t)
  =\int_{\bbS^{2}}f(\omegaNu,\epsilonNu,\bx,t)\,\big\{1,l^{i},l^{i}\,l^{j}\}\,d\omegaNu.  
  \label{eq:angularMoments}
\end{equation}
Here, $\{l^{i}\}$ are components of a unit three-vector proportional to the neutrino three-momentum $\bp=\epsilonNu\,\bl$.  
(We adopt the Einstein summation convention and let repeated Latin indices run from $1$ to $3$.)
In a kinetic description, neutrinos are described by the phase space distribution function $f$, and the momentum space coordinates used here are the neutrino energy $\epsilonNu$, and the angles $\thetaNu$ and $\phiNu$ (describing the neutrino propagation direction relative to a local orthonormal basis).  
In forming the angular moments in Eq.~\eqref{eq:angularMoments}, the integrals extend over the unit sphere
\begin{equation}
  \bbS^{2}=\big\{\,(\thetaNu,\phiNu)~|~\thetaNu\in[0,\pi),\,\phiNu\in[0,2\pi)\,\big\}.  
\end{equation}
The angular moments are related to the neutrino energy density, momentum density, and stress ($\cJ$, $\cH^{i}$, and $\cK^{ij}$, respectively).  
They are functions of the neutrino energy $\epsilonNu$, spatial position $\{x^{i}\}$, and time $t$.

To model neutrino matter interactions, we include emission and absorption, (isotropic) elastic scattering on nucleons and nuclei, and neutrino-electron scattering (NES).  
On the right-hand sides of Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, $\tilde{\chi}(\epsilonNu,\bU)$ and $\kappa(\epsilonNu,\bU)=\tilde{\chi}(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$ are absorption and transport (absorption plus elastic scattering) opacities, respectively.  
The absorption opacity $\tilde{\chi}$ is corrected for stimulated absorption \citep[cf.][]{mezzacappaBruenn_1993a}.  
The NES rates $R^{\In}$ and $R^{\Out}$ are obtained by expanding the full kernels in a Legendre series, retaining only the isotropic part \citep[see, e.g.,][]{cernohorsky_1994}, and the integrals extend over all energies $\bbR^{+}=\{\epsilonNu~|~\epsilonNu\in[0,\infty)\}$.  
The energy volume element is $dV_{\epsilonNu}=\epsilonNu^{2}d\epsilonNu$.  
The opacities depend on the neutrino energy and the local thermodynamic state of the fluid $\bU=(\rho,T,Y_{e})^{T}$, where $\rho$ is the (baryon) mass density, $T$ the temperature, and $Y_{e}$ the electron fraction.  
For simplicity we consider a single neutrino species (electron neutrinos), and we use opacities given by \citet{bruenn_1985}; see Section~\ref{sec:opacities} for further details.  
In Eq.~\eqref{eq:momentEquationEnergy}, the equilibrium density $\cJ_{0}=4\pi\,f_{0}(\epsilonNu,\bU)$ is computed from the angular moment of the isotropic Fermi-Dirac distribution
\begin{equation}
  f_{0}(\epsilonNu,\bU)=\f{1}{e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T}+1},
\end{equation}
where $k_{\mbox{\tiny B}}$ is Boltzmann's constant, and $\mu_{\nu}=\mu_{p}-\mu_{n}+\mu_{e}$ is the neutrino chemical potential.  
(The chemical potentials of protons, neutrons, and electrons are denoted by $\mu_{p}$, $\mu_{n}$, and $\mu_{e}$, respectively.)

>>>>

\documentclass[11pt,letterpaper,twoside,english,final]{article}

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{mathrsfs}
\usepackage{definitions}

\newcommand{\NES}{\mbox{\tiny NES}}
\newcommand{\dx}{\Delta x}
\newcommand{\sumx}{\sum_{i=1}^{d}}
\newcommand{\thornado}{\texttt{thornado}}

\input{ornlmacro}

@@

%\bibliographystyle{plain} 
\begin{titlepage}

%NOTICE about prelim document
%delete NOTICE when document become final
%\vspace{-1.75in}
%\hspace{-1in}
%\par
%\footnotesize
%\vspace{-.8in} 
%  \fbox{
%  \begin{minipage}{0.55\textwidth}
%   \raggedright
% \textsf{\textbf{NOTICE: This document contains information of a preliminary nature and is not intended for release.  It is subject to revision or correction and therefore does not represent a final report.}}
%  \end{minipage}
%  }

\URCornerWallPaper{1.0}{Graybar.png}
\ThisURCornerWallPaper{1.0}{Graybar.png} 
\ClearWallPaper

\begin{flushright}{\textsf{\bfseries{\large{ORNL/TM-2017/501}}}}\\ 
\end{flushright}

\vspace{.15in}

\noindent
\begin{shaded}
\Huge{\textsf{\bfseries\color{white}{TOWARDS A DISCONTINUOUS GALERKIN METHOD FOR THE MULTI-GROUP TWO-MOMENT MODEL OF NEUTRINO TRANSPORT}}}
\end{shaded}

\begin{figure}[H]
\flushright
\includegraphics[width=4.88in, height=3.47in]{Officiallogo.png} 
\end{figure}

\def\author#1{\large{\textsf{#1}}\\ }

\def\date#1{{\vspace{.25 in}

\large{\bfseries{\textsf{#1}}}\\ }}

\def\authordate#1{

\noindent
\vskip 6.8pt
\vtop to0pt{
\setlength{\parindent}{112mm}

#1\vss}}

\authordate{
\author{Eirik Endeve}
\author{Ran Chu}
\author{Anthony Mezzacappa}
\author{Bronson Messer}
\date{\today}}

% delete NOTICE when document become final
%\DraftBox                    % command for Draft. Not for public release.
%\NotPublic                % command for Not for public release.
\ApprovedBox             % command for Approved for public release.
%\BusSenBox                 % command for Business Sensitive.
%\OUOBox                    % command for Official Use Only. To add item, open ORNLmacro.tex, make change, and save
%\ECIBox                      % command for Export Controlled Information
%\OUOECIBox               % command for OUO and ECI
%\OUOATBox                % command for Official Use Only and Applied Technology
%\OUOECIATBox           % command for Official Use Only, Export Controlled Information, and Applied Technology
%\UCNIBox                   % command for Unclassified Controlled Nuclear Information
%\CRADABox                % command for CRADA final.

\setlength{\parindent}{0pt}

\bigskip
\noindent

\thispagestyle{fancy}
\fancyfoot{}
\fancyfoot[L]{\includegraphics[width=3.72in]{textlogo.png}}
\renewcommand{\headrulewidth}{0pt}
\addtolength{\footskip}{0.3in}

\end{titlepage}

\newpage

 \newpage

\thispagestyle{empty}                           % Hide when using other pagestyles (Not for public release, OUO, etc.)
%\thispagestyle{mypagestylepublic}        % Use for "Not for public release." in footer with no page numbers
%\thispagestyle{mypagestyleempty}    % Use for "OUO" in footer with no page numbers 
\vspace*{\fill}

  \par

  \par
  \begin{center}
    \fbox{
      \parbox{5.61in}{
        \small
        \begin{center}

         {\bf\textsf{ DOCUMENT AVAILABILITY}}
        \end{center}
    \textsf{Reports produced after January 1, 1996, are generally available free via US Department of Energy (DOE) SciTech Connect. }
          \\
          ~

          \hspace{0.5in}{\textsf{\textbf{\emph{ Website:}}}} \textsf{\url{ http://www.osti.gov/scitech/ }}
          \\
          ~
          \\
            \textsf{Reports produced before January 1, 1996, may be purchased by members of the public from the following source:}
          \\  
          ~

          \hspace{0.5in}   \textsf{National Technical Information Service }

          \hspace{0.5in}   \textsf{5285 Port Royal Road }

          \hspace{0.5in}   \textsf{Springfield, VA 22161 }

          \hspace{0.5in} {\textsf{\textbf{\emph{Telephone:}}}}   \textsf{703-605-6000  (1-800-553-6847) }

          \hspace{0.5in} {\textsf{\textbf{\emph{TDD:}}}}   \textsf{703-487-4639 }

          \hspace{0.5in} {\textsf{\textbf{\emph{Fax:}}}}   \textsf{703-605-6900 }

          \hspace{0.5in} {\textsf{\textbf{\emph{E-mail:}}}}   \textsf{\href{mailto:info@ntis.gov}{info@ntis.gov}}

          \hspace{0.5in} {\textsf{\textbf{\emph{Website:}}}}   \textsf{\url{http://classic.ntis.gov/}} 
          \\
          ~
          \\
          \textsf{Reports are available to DOE employees, DOE contractors, Energy Technology Data Exchange representatives, and International Nuclear Information System representatives from the following source: }
          \\
          ~

          \hspace{0.5in}\textsf{Office of Scientific and Technical Information}

          \hspace{0.5in}\textsf{PO Box 62}

          \hspace{0.5in}\textsf{Oak Ridge, TN  37831}

          \hspace{0.5in}{\textsf{\textbf{\emph{Telephone:}}}}\textsf{ 865-576-8401}

          \hspace{0.5in}{\textsf{\textbf{\emph{Fax:}}}}\textsf{ 865-576-5728}

          \hspace{0.5in}{\textsf{\textbf{\emph{E-mail:}}}} \textsf{\href{mailto:reports@osti.gov}{report@osti.gov}}

          \hspace{0.5in}{\textsf{\textbf{\emph{Website:}}}} \textsf{\url{http://www.osti.gov/contact.html}}

         }
      }
  \end{center}
  \par
  \vspace{0.1in}
  \par
  \begin{center}
   \small
    \fbox{\parbox{4in}{

       \textsf{This report was prepared as an account of work sponsored by an agency of the United States Government. Neither the United States Government nor any agency thereof, nor any of their employees, makes any warranty, express or implied, or assumes any legal liability or responsibility for the accuracy, completeness, or usefulness of any information, apparatus, product, or process disclosed, or represents that its use would not infringe privately owned rights. Reference herein to any specific commercial product, process, or service by trade name, trademark, manufacturer, or otherwise, does not necessarily constitute or imply its endorsement, recommendation, or favoring by the United States Government or any agency thereof. The views and opinions of authors expressed herein do not necessarily state or reflect those of the United States Government or any agency thereof. }}}

  \end{center}

\vspace*{\fill}

\newpage

\begin{titlepage}
%\thispagestyle{mypagestylepublic}        % Use for "Not for public release." in footer with no page numbers
%\thispagestyle{mypagestyleempty}    % Use for "OUO" in footer with no page numbers

\begin{flushright}{\textsf{\bfseries{\large{ORNL/TM-2017/501}}}}\\ 
\end{flushright}

\vspace{0.5in}

\begin{center}
Computer Science and Mathematics Division
\end{center}

\vspace{1.25in}

\begin{center}
{\bf{\large{TOWARDS A DISCONTINUOUS GALERKIN METHOD FOR THE MULTI-GROUP TWO-MOMENT MODEL OF NEUTRINO TRANSPORT}}}\\
\vspace{0.5in}

Eirik Endeve, Ran Chu, Anthony Mezzacappa, Bronson Messer

\vspace{1.35in}

Date Published: January 2018

\vspace{1.35in}

Prepared by \\
OAK RIDGE NATIONAL LABORATORY \\
Oak Ridge, TN 37831-6283 \\
managed by \\
UT-Battelle, LLC \\
for the \\
US DEPARTMENT OF ENERGY \\
under contract DE-AC05-00OR22725

\end{center}

\end{titlepage}

\clearemptydoublepage 
\begin{centering}
\tableofcontents 
\end{centering}

%\pagestyle{mypagestylepublicpage}      % Use for "Not for public release." in footer with page numbers
%\thispagestyle{mypagestyle}                  % Use for "OUO" in footer with page numbers

\pagenumbering{roman}

\setcounter{page}{3}

\newpage

\clearemptydoublepage
\phantomsection 
\begin{centering}
\listoffigures 
\end{centering}

\addcontentsline{toc}{section}{LIST OF FIGURES}

\newpage

\clearemptydoublepage
\phantomsection 
\begin{centering}
\listoftables 
\end{centering}

\addcontentsline{toc}{section}{LIST OF TABLES}

\newpage

%\pagestyle{mypagestylepublicpage}      % Use for "Not for public release." in footer with page numbers
%\thispagestyle{mypagestyle}                  % Use for "OUO" in footer with page numbers

\clearemptydoublepage
\phantomsection
\addcontentsline{toc}{section}{ACRONYMS} 
\begin{center}
{\bf{ACRONYMS}}
\end{center}

\begin{table}[h]
\vspace{-6pt}
\begin{tabular}{l l} 
ORNL & Oak Ridge National Laboratory \\
CCSN & Core-Collapse Supernova \\
PNS & Proto-Neutron Star \\
NES & Neutrino-Electron Scattering \\
SI & Semi-Implicit \\
IMEX & Implicit-Explicit \\
SIRK & Semi-Implicit Runge-Kutta \\
DG & Discontinuous Galerkin \\
LF & Lax-Friedrichs \\
HLL & Harten-Lax-van Leer \\
PDE & Partial Differential Equation \\
ODE & Ordinary Differential Equation \\
CFL & Courant-Friedrichs-Lewy \\
RP & Realizability-Preserving \\
SSP-RK & Strong Stability-Preserving Runge-Kutta \\
\end{tabular}
\end{table}

\newpage

\clearemptydoublepage
\phantomsection
\addcontentsline{toc}{section}{ADDITIONAL FRONT MATERIAL}
\begin{center}
{\bf{ADDITIONAL FRONT MATERIAL}}
\end{center}

%Material after the Contents, List of Figures, List of Tables, and the Acronym list may include any or all of the following sections and should begin on an odd-numbered page in the order listed below:
%
%\noindent{FOREWORD {\color{blue}{\emph {[Note: Spelling is not ``FORWARD.'']}}}\\
%\noindent{PREFACE}\\

\noindent{ACKNOWLEDGMENTS}\\

This research used resources of the Oak Ridge Leadership Computing Facility, which is a DOE Office of Science User Facility supported under Contract DE-AC05-00OR22725.  
This research was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration.  
A portion of this research used resources at the Joint Institute for Computational Sciences (JICS), a DOE Office of Science User Facility operated by the Oak Ridge National Laboratory.

thornado{\_}mini uses equation of state and opacity tables and subroutines developed under the WeakLib project.  
We acknowledge the contributions to WeakLib of W.~Raphael Hix, Ryan Landfield, and Eric J. Lentz.

%\noindent{EXECUTIVE SUMMARY OR SUMMARY}\\
%\noindent{ABSTRACT}
%
%Each section of the front material begins on an odd-numbered page. The abstract, if very brief, may begin on page 1 inserted above the introduction section.
%
%\def\thefootnote{\fnsymbol{ctr}} 
%\long\def\symbolfootnote[#1]#2{\begingroup
%\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}
%
%\bigskip
%\phantomsection
%\addcontentsline{toc}{section}{GENERAL INFORMATION}
%\begin{center}
%{\bf{GENERAL INFORMATION}}
%\end{center}
%
%Document margins are 1 in. all around.
%
%Text footnotes are 9 pt., TNR with no space between notes.\symbolfootnote[1]{Footnotes are 9pt., TNR with no space between notes. Footnotes are 9 pt., TNR with no space between notes.} Footnote indicators in text are superscript symbols and should appear in the following order: $\ast$, $\dagger$, $\ddagger$, $\S$, $\ast\ast$, $\dagger\dagger$, $\ddagger\ddagger$. Symbols {\underline {are not}} used as table footnote indicators (see page 4).
%
%Footnotes may be used in addition to the author-date citation method, in which case, the footnotes provide additional information to the text, not bibliographic citations.

\newpage

\clearemptydoublepage

\pagenumbering{arabic}

\setcounter{page}{1}

\phantomsection
\addcontentsline{toc}{section}{ABSTRACT}
\begin{center}
{\bf{ABSTRACT}}
\end{center}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

We present progress on the development of a Discontinuous Galerkin (DG) method for solving the non-relativistic, multi-group two-moment model for neutrino transport.  
The method --- ultimately targeted at simulation of neutrino transport in core-collapse supernovae --- includes the use of curvilinear coordinates and tabulated microphysics (i.e., nuclear equation of state and weak interaction rates) to model neutrino-matter coupling.  
In this report, we provide algorithm details and present results from an extensive set of benchmark problems in one spatial dimension. 
The algorithms have been implemented in the Toolkit for High-Order Neutrino-Radiation Hydrodynamics (\thornado), and Fortran code used to generate the results has been made available to the public through GitHub.\footnote{https://github.com/ECP-Astro/thornado{\_}mini}

\begin{center}
  \section{INTRODUCTION}
\end{center}

The explosion of a massive star (i.e., with mass $\gtrsim8$ solar masses) in a core-collapse supernova (CCSN) is initiated by the collapse of its iron or oxygen-neon-magnesium core, the product of millions of years of stellar evolution. 
The collapse proceeds to densities exceeding the density of nucleons in the atomic nucleus, at which point the inner core becomes incompressible, ``rebounds," and launches a shock wave into the outer stellar core. 
This shock wave ultimately propagates through the outer layers of the star and disrupts it in a CCSN explosion. 
However, before this, the shock wave stalls in the outer core, due to energy losses from neutrino emission and dissociation of heavy nuclei, and turns into an accretion shock.  
Exactly how the shock wave is revived remains a central question in CCSN theory.  
After core bounce, about $10^{53}$~ergs of energy in the form of neutrinos and antineutrinos of all three flavors (i.e., electron, muon, and tau neutrinos) is emitted from the proto-neutron star (PNS) at the center of the explosion.  
Early simulations \citep[e.g.,][]{betheWilson_1985} have shown that energy from these neutrinos can be deposited behind the shock and revive it, and this neutrino-reheating mechanism is central to current CCSN theory.  
See, e.g., \citet{mezzacappa_2005,janka_2012,burrows_2013,muller_2016}, for comprehensive reviews.

Today, simulations play an indispensable role in disentangling details of the CCSN explosion mechanism, and in connecting observables to physical processes in the explosion.  
Observables include nucleosynthetic yields, and optical and neutrino signals.  
Simulations may also play a role in the future detection and interpretation of gravitational wave signals.  
As such, CCSNe are multi-messengers, and CCSN models are quintessentially multi-physics --- involving all the known fundamental forces of nature.  
However, the computational cost of CCSNe models is dominated by solving equations describing the transport of neutrinos and their interaction with the stellar fluid. 
Neutrinos diffuse away from the PNS, become semi-transparent, and eventually free-streaming --- carrying information about the physical conditions in the exploding star.  
One of the principal goals of a CCSN model is to compute the lepton and four-momentum exchange between the neutrinos and the stellar fluid.

Net energy transfer from neutrinos to the stellar fluid occurs in the layer between the so-called \emph{gain radius} and the shock --- the \emph{heating region}.  
Importantly, in the heating region, the neutrino mean free path exceeds other characteristic length scales (the Knudsen number exceeds unity), and non-equilibrium effects may become important.  
To capture non-equilibrium effects, a \emph{kinetic description} is required, where the neutrinos are described by the \emph{distribution function} $f(\omegaNu,\epsilonNu,\vect{x},t)$, a phase space density governed by the Boltzmann equation.  
The distribution function gives, at time $t$, the number of neutrinos at spatial position $\vect{x}\in\mathbb{R}^{3}$, with energy $\epsilonNu\in\mathbb{R}^{+}$, propagating in direction $\omegaNu\in\mathbb{S}^{2}$ --- a seven-dimensional function.

In this work, we seek to develop efficient methods to model neutrino transport in CCSN simulations.  
On the one hand, the need for high spatial resolution and unconstrained spatial dimensionality to follow important fluid instabilities as the explosion develops makes direct solution of the Boltzmann equation a formidable computational challenge.  
On the other hand, neutrino heating rates are particularly sensitive to the neutrino energy distribution, which requires retention of the energy dimension of momentum space.  
Therefore, to balance manageable computational cost with required physical fidelity, the number of unknowns can be reduced by a truncated moment expansion of the angular dimensions of momentum space; i.e., by solving for a finite number of angular moments of the distribution function
\begin{equation}
  \vect{\mathcal{M}}(\epsilonNu,\vect{x},t)=
  \int_{\mathbb{S}^{2}}f(\omegaNu,\epsilonNu,\vect{x},t)\,\vect{w}(\omegaNu)\,d\omegaNu,
\end{equation}
where $\vect{w}$ are momentum space angular weighting functions.  
The two-moment model \citep[e.g.,][]{andersonSpiegel_1972} is obtained by truncating the moment expansion at the level of the so-called first order moments (akin to an expansion in spherical harmonics up to degree $\ell=1$), so that the unknown moments are the spectral particle density $\cJ$ and particle flux $\bcH$ --- a two-moment model; $\bcM=(\cJ,\bcH)^{T}$.  
This is the basic system we will consider.  
Note that the moments depend on the neutrino energy, spatial position, and time, and are governed by a system of nonlinear moment equations.

The solution to a truncated moment model is confronted with the challenge of specifying higher order moments (e.g., radiation stresses), which appear in the governing moment equations.  
In this work, the solution to this \emph{closure problem} is approximated by relating these higher order moments to the evolved moments through relatively simple analytic expressions \citep[cf.][]{minerbo_1978,levermore_1984,cernohorskyBludman_1994}.  
Obviously, this approach limits the fidelity of the model when the angular distribution of the radiation field is highly anisotropic.  
However, this shortcoming can (at additional computational cost) be alleviated by extending the moment expansion to include more moments, or by solving the Boltzmann equation directly, in regions where higher fidelity is needed.  
In this approach, the two-moment model becomes a component in a multi-model hierarchy, and the developments described here can be fully leveraged.  
We note that this multi-model approach introduces additional challenges associated with model coupling.

The neutrino-radiation hydrodynamics system supports three important time scales: (i) the \emph{radiative time scale} $\tau_{\mbox{\tiny r}}$, governed by the speed of light; (ii) the \emph{fluid time scale} $\tau_{\mbox{\tiny f}}$, governed by the fluid flow and sound speeds; and (iii) the \emph{collision time scale} $\tau_{\mbox{\tiny c}}$, governed by the mean time between neutrino-matter interactions.  
In the CCSN environment, the flow \emph{and} sound speeds approach significant fractions of the speed of light, while $\tau_{\mbox{\tiny c}}$ is very short in the dense PNS.  
Thus, we have the time scale ordering: $\tau_{\mbox{\tiny c}}\ll\tau_{\mbox{\tiny r}}\lesssim\tau_{\mbox{\tiny f}}$.  
We will not resolve the short timescales imposed by collisions in the PNS, but integrate the stiff collision term with \emph{implicit methods}.  
Streaming terms will be integrated in time using \emph{explicit methods}.  
This motivates the use of \emph{Semi-Implicit (SI)}, or \emph{Implicit-Explcit (IMEX)}, time integration methods (e.g., \cite{minion_2003,pareschiRusso_2005,chertock_etal_2015}) for optimal balance between stability and efficiency.  
Specifically, we will use Semi-Implicit Runge-Kutta (SIRK) methods, similar to those discussed by \citet{chertock_etal_2015}.  
This combination of explicit and implicit methods, as opposed to fully implicit methods, avoids  the need to solve large, distributed systems of algebraic equations for each time step.

To solve the neutrino transport equations we discretize the moment equations in energy-position space using high-order Discontinuous Galerkin (DG) methods (e.g., \cite{cockburnShu_2001,hesthavenWarburton_2008}).  
DG methods combine elements from both spectral and finite volume methods, and are an attractive option for solving hyperbolic partial differential equations (PDEs).  
They achieve high-order accuracy on a compact stencil; i.e., data is only communicated with nearest neighbors, regardless of the formal order of accuracy.  
This leads to a high computation to communication ratio, and favorable parallel scalability on heterogeneous architectures has been demonstrated \citep{klockner_etal_2009}.  
DG methods can be used in combination with $hp$-adaptivity \citep{remacle_etal_2003}, where, in addition to grid refinement with AMR, the local polynomial degree can be chosen differently, and independently, in different cells.  
They can easily be applied to problems involving curvilinear coordinates, which is beneficial in numerical relativity \citep{teukolsky_2016}.  
Importantly, DG methods, as opposed to, e.g., finite volume methods, exhibit favorable properties when collisions are included.  
Specifically, without modification, they recover the correct asymptotic behavior in the diffusion limit, characterized by frequent collisions \citep[e.g.,][]{larsenMorel_1989,adams_2001,guermondKanschat_2010}.  
The DG method was introduced in the 1970s by \citet{reedHill_1973} to solve the neutron transport equation, and has undergone remarkable developments since then \citep[see, e.g.,][and references therein]{shu_2016}.  
It has only recently received attention from the computational astrophysics community \citep[e.g.,][]{radiceRezzolla_2011,schaal_etal_2015,wuTang_2017}, mainly to solve the Euler equations for ideal hydrodynamics, but also to solve Einstein's field equations \citep{teukolsky_2016}.  
DG methods have also been considered for the neutrino transport problem in CCSN \citep[e.g.,][]{radice_etal_2013,endeve_etal_2015}.

The ultimate goal of this work is to develop genuinely multi-dimensional, high-order numerical methods within the DG framework for solving the equations of neutrino-radiation hydrodynamics, that include relativity and full nonlinear coupling to the stellar fluid through a complete set of weak interactions, to enable high fidelity simulations of CCSN explosions.  
This is a formidable challenge, and, in our opinion, this goal is best achieved in incremental steps.  
The work reported here is the initial step toward this goal, where we present developments of a DG method for solving the non-relativistic, multi-group two-moment model in one spatial dimension (but sufficiently general to accommodate Cartesian, spherical, and cylindrical spatial coordinates).  
We have tabulated a basic set of neutrino-matter interactions (electron capture and elastic scattering on nucleons and nuclei, and neutrino-electron scattering, as described by \citet{bruenn_1985}), using a tabulated nuclear equation of state (EoS), and we include lepton and energy exchange with the stellar fluid.  
The algorithms have been implemented in the Toolkit for High-Order Neutrino-Radiation Hydrodynamics (\thornado), and we provide a limited public release of the algorithms on GitHub (\url{https://github.com/ECP-Astro/thornado_mini}).

This report is organized as follows:  
We describe the non-relativistic, multi-group two-moment model we consider in Section~\ref{sec:twoMomentModel}
We detail the spatial discretization of the two-moment model using the DG method in Section~\ref{sec:dgMethod}  
The time integration methods used for problems with and without collisions are described in Section~\ref{sec:timeIntegration}  
Limiting techniques to control nonphysical oscillations and to prevent nonphysical states are described in Section~\ref{sec:limiters}
We provide details on the tabulated opacities in Section~\ref{sec:opacities}  
Results from an extensive set of benchmark problems (in Cartesian, spherical, and cylindrical spatial coordinates), covering streaming and collision dominated limits, but with analytic opacities are presented in Section~\ref{sec:numericalBasic}  
The neutrino-matter coupling method and results from two tests using tabulated opacities are discussed in Section~\ref{sec:numericalTableOpacities} 
Finally, we summarize and discuss next steps in Section~\ref{sec:summary}

\begin{center}
  \item\section{TWO-MOMENT MODEL FOR NEUTRINO TRANSPORT}
  \label{sec:twoMomentModel}
\end{center}

Here we give a description of the two-moment model to be solved with the DG method.

\subsection{BASIC MATHEMATICAL MODEL}
